Characterization and Photometric Performance of the Hyper Suprime-Cam Software Pipeline.

### 2017-04-27 draft/ms_v5.tex
Synthetic data

    \hscpipe{} is a complex data reduction pipeline that still faces many challenges
    and is under active development.
    Unlike most photometric pipelines on the market, it involves high-level reduction
    processes that produce not only deep \coadd{} images but also a series of
    science-ready catalogs. 
    To perform tests of the data products that result from \hscpipe{}, \synpipe{} is 
    designed to satisfies a few basic standards:

@@

        \item \textbf{Flexible Capabilities}: 
            \synpipe{} is designed to be flexible enough to be useful for HSC users 
            with a range of scientific goals.
            We provide default catalogs to generate samples of synthetic stars
            and galaxies with realistic magnitudes and color distributions along with
            tools to help the user work on HSC DR1 data. 
            However, \synpipe{} also easily allows users to design and use their own 
            custom made input catalogs.
            \synpipe{} is already being used in a wide range of
            scientific applications (see Section \ref{sec:summary} for details).

@@

        \item \textbf{Input catalog}. 
            This is a catalog in \texttt{FITS} (Flexible Image Transport System)
            format that contains the coordinates, magnitudes, and model parameters of 
            synthetic objects.  
            The positions of synthetic galaxies can be specified via the input catalog, 
            they can be distributed randomly over a given region, or on a grid 
            (the grid option is useful in that it avoids blends between Synthetic 
            objects, e.g. see Murata\etal in prep.).

@@

        \item With the input catalog and test data information in hand, \synpipe{} 
            injects synthetic objects into single-exposure images (step
            \texttt{addFakes.py}; see the middle panel of Fig \ref{fig:flowchart}).
            \synpipe{} goes through all the individual CCD images that belong to 
            each \visit{} and decides which synthetic objects from the input catalog 
            need to be injected. 
            \synpipe{} uses the \textit{initial astrometric calibration} of each 
            single \visit{} to convert the input coordinates into locations on the 
            CCD image. 
            For each synthetic object, \synpipe{} uses the reconstructed PSF for each 
            exposure.  
            The photometric zero point from the single-\visit{} calibration is used
            to convert input magnitudes of synthetic objects into fluxes.

@@

%% ------------------------------------------------------------------------------------ %%            
        If necessary, an additional shear ($g_1$ and $g_2$) can be applied to the 
        model at this point.
        Then, \synpipe{} passes the reconstructed PSF image into a
        \galsim{}\texttt{InterpolatedImage} object and convolves it with the galaxy 
        model using \galsim{}\texttt{.Convolve}.
        After this, \synpipe{} uses the \galsim{}\texttt{.drawImage} method to create 
        the image of the simulated galaxy.
        A component with a high \ser{} index ($n_{\mathrm{Ser}} > 4$) often requires 
        a large image size to cover all of its flux, so \synpipe{} allows the user 
        to truncate the model at a given radius (N$\times R_{\mathrm{e}}$).
        \synpipe{} will also give warning information when a component with $
        n_{\mathrm{Ser}} > 6$ is encountered because it takes much longer to achieve
        accurate PSF convolution in such a model; and the result is, therefore,
        not a realistic galaxy model.

        \item After \synpipe{} creates the image of the simulated galaxy, it then 
            shifts the image according to is location on the CCD on a pixel-by-pixel 
            basis; \synpipe{} also crops the images when necessary.
            When that task is complete, \synpipe{} adds appropriate noise to the 
            image based on the pixels it covers and the calibration information of 
            the detector, and creates a variance image that reflects the influence 
            of the synthetic object.
            Since each HSC CCD consists of 4 amplifiers, each with different 
            characteristics, \synpipe{} carefully creates a corresponding gain map 
            to provide accurate noise level.

@@

    Using the infrastructures provided by \hscpipe{}, the user can easily perform these 
    steps in parallel, and the overall efficiency of \synpipe{} is similar to
    the real data reduction process. 
    The data and catalog volumes is also similar to the original data reduction process.

    The final step is to identify Synthetic galaxies in the output catalogs. 
    \synpipe{} can match the input catalog to the output catalog (this contains a mix 
    of real and synthetic galaxies) using a matching radius that is specified by the 
    user.
    Each unmatched object from the input catalog is also passed to the results catalog
    with a unique label.
    In the case of multiple matches, the user can choose between returning only the
    closest match or returning all objects within the matching radius (the nearest one
    is labeled as such).
%% ------------------------------------------------------------------------------------ %%

@@

%% ------------------------------------------------------------------------------------ %%
\subsection{Data}
    We use \tract{}$=9699$ in the VVDS field (median FWHM$=0.449$\asec{} in $i$-band;
    in this paper referred to as \texttt{goodSeeing}) and
    \tract{}$=8764$ in the XMM-LSS field (median FWHM$=0.700$\asec{};
    \texttt{badSeeing}).
    \tract{}$=9699$ also has better seeing in both $r$- and $z$-bands;
    the seeing conditions in $g$- and $y$-bands are very similar for both
    \tract{}$=9699$ and \tract{}$=8764$.
    For both \tracts{}, the \coadd{} image in each band includes data from
    20--40 \visits{}.  
    These two \tract{} are selected because they are not at the edge of a field, do 
    not contain any extremely bright ($i<12$ mag) saturated stars, and are 
    representative of both "good" and "bad" seeing conditions in the $i$-band.
%% ------------------------------------------------------------------------------------ %%

@@

    We use a synthetic star sample built using data from the HST/ACS catalog of 
    \citet{Leauthaud2007}.  
    We first select stars from the \citet{Leauthaud2007} catalog (this classification 
    is reliable down to $I_{\mathrm{F814W}}{\sim}25.2$ mag). 
    We then match this sample against our HSC UltraDeep--COSMOS data (which reaches 
    ${\sim}27.2$ mag in $i$-band). 
    After applying basic quality cuts (see \ref{app:qc}), this procedure provides us 
    with five-band HSC photometry for 14,472 stars.
    To increase the sample size, we re-sample the five-band magnitude distributions 
    using the \texttt{astroML} (\citealt{astroml}) implementation of the 
    \texttt{extreme-deconvolution} algorithm developed by \citet{Bovy2011}
    \footnote{\texttt{https://github.com/jobovy/extreme-deconvolution/}} to
    generate 100,000 synthetic stars. 
    Fig \ref{fig:star_sample} shows the magnitude and color distributions of real 
    COSMOS stars compared to the resampled distributions in all five-bands and in 
    color-color space.

    For synthetic galaxies, we use single-\ser{} galaxy models of COSMOS galaxies with 
    $I_{F814W} \leq 25.2$ mag. 
    These are models from \citet{Lackner2012} applied to \textit{HST}/ACS images and 
    are included in \galsim{}. 
    Each model is described by a total of five parameters: magnitude, effective radius,
    \ser{} index, axis ratio, and position angle.
    Although the single-\ser{} model is not always the best choice for describing 
    galaxies, its flexibility enables it to reasonably describe the flux distribution 
    of most galaxies.
    Also, its simplicity makes it easy for us to diagnose potential problems with the 
    \cmodel{} photometry.
    We exclude a tiny fraction (${\sim}4$\%) of ill-behaved models that have a very 
    high \ser{} index ($n_{\mathrm{Ser}} > 6.0$) or a very low central surface 
    brightness ($\mu_{i} < 24.0$\sb).
    We further match this COSMOS sample with the HSC UltraDeep--COSMOS data.
    After removing objects with problematic photometry, we obtain a sample of
    58,210 synthetic galaxies with realistic five-band HSC photometry.
    As shown in Fig \ref{fig:galaxy_sample}, the majority of these galaxies are
    faint ($i<24.0$ mag) and barely resolved ($R_{\mathrm{e}}< 1.0$\asec).
    This sample is appropriate to test \hscpipe{}'s general photometric behaviors, 
    but the sample lacks relatively bright galaxies ($i<20.5$ mag).

    We spatially distribute synthetic stars and galaxies randomly in our two selected 
    \tracts{}. 
    For stars, we use a number density of 1,000 per CCD. 
    For galaxies, we use a number density of 500 per CCD.

@@

    We use HSC DR1 data stored at Kavli Institute for the Physics and Mathematics of 
    the Universe (Kavli IPMU). 
    For stars the \texttt{addFakes.py} step takes ${\sim}1.5$ hours per \tract{} in 
    each band. 
    The same process takes longer for galaxy tests (${\sim}3.0$ hours per
    \tract{} in each band) because the \galsim{} simulation is more time consuming.
    The \texttt{stack.py} step and \texttt{multiBand.py} step together take 
    ${\sim}3.5$ hours per \tract{}.

    We match the results with the input catalogs using a 2--pixel matching radius
    to generate output catalogs for \forced{} and \unforced{} photometry
    in each band. For our tests, we reject  synthetic objects  locate within 2 pixels 
    of a real HSC object.

@@

    Here we discuss \hscpipe{}'s general performance, assessed mainly through \forced{} 
    PSF photometry (for stars) and (\cmodel{}) photometry (for galaxies).
    Although \hscpipe{} does provide other options (e.g., aperture and Kron photometry),
    these two are the only options that consider the effects of PSF in different bands 
    and are consistent across all bands in terms of position and shape.

@@

    In \hscpipe{}, the PSF magnitude is derived using a matched-filter method that
    depends on the best-fit PSF model and uses the Lanczos interpolation to shift
    the PSF model.
    The error of the PSF magnitude from \hscpipe{} only considers the per-pixel noise,
    and does not include centroid uncertainties.
    \hscpipe{} also estimates aperture correction for PSF photometry.
    For more details please see Bosch\etal (in prep.).

@@

    Next, we exclude stars that are misidentified by \hscpipe{} as extended
    objects using the \texttt{classification.extendedness} parameter in each band.
    The fraction of misclassification is ${\sim}10$--$20$\%, and clearly depends on
    seeing conditions.
    For the same reason, $g$-band has the highest fraction of misclassification,
    while $i$-band is recommended for selecting point sources.
    The star--galaxy separation issue is discussed more in Section \ref{ssec:sg}.

    In the following comparisons, we show the results from the
    \textbf{good-seeing} and \textbf{bad-seeing} \tracts{} separately because it
    is important to understand how the seeing condition affects photometric accuracy 
    as well as to test whether \hscpipe{} can deliver unbiased photometry under 
    different seeing conditions.

    Besides seeing, the degree to which an object is blended with other objects 
    is another factor that influences photometric accuracy. 
    To quantify the impact of blending effects, we use the ``blendedness'' 
    parameter, $b$. 
    This parameter describes how much any given object overlaps with neighboring 
    objects (see Appendix \ref{app:defineb} and Murata\etal in prep.). 
    We divide our sample according to the blendedness parameter. 
    A value of $b=0$ corresponds to isolated objects whereas objects with $b=1$ 
    complete overlap with other objects. 
    Here we use $b>0.05$ to define highly blended stars (${\sim}4$--$9$\% are blended).
    The fraction of highly blended stars slightly increases in the bad seeing \tract{}. 
    The impact of blendedness will be discussed more in Section \ref{ssec:blendedness}.

@@

    %% C %%
    \bibitem[Chang et al.(2015)]{Chang2015} Chang, C., Busha, M.~T., Wechsler, R.~H.,
             et al.\ 2015, \apj, 801, 73

    %% D %%
    \bibitem[Dressler et al.(2012)]{Dressler2012} Dressler, A., Spergel, D., Mountain,
             M., et al.\ 2012, arXiv:1210.7809

@@

    \bibitem[Calabretta \& Greisen(2002b)]{WCS2} Calabretta, M.~R., \& Greisen, E.~W.\
             2002, \aap, 395, 1077

>>>>

    \hscpipe{} is a complex data reduction pipeline that still faces many challenges
    and is under active development.
    Unlike most publicly available photometric pipelines, it involves high-level 
    reduction processes that produce not only deep \coadd{} images but also a series 
    of science-ready catalogs. 
    To perform tests of the data products that result from \hscpipe{}, \synpipe{} is 
    designed to satisfies a few basic standards:

@@

        \item \textbf{Flexible Capabilities}: 
            \synpipe{} is designed to be flexible enough to be useful for HSC users 
            with a range of scientific goals.
            We provide default catalogs to generate samples of synthetic stars
            and galaxies with realistic magnitudes and color distributions along with
            tools to help the user work on HSC DR1 data. 
            However, the users can also supply their own input catalogs best suited 
            for their applications.
            \synpipe{} is already being used in a wide range of
            scientific topics (see Section \ref{sec:summary} for details).

@@

        \item \textbf{Input catalog}. 
            This is a catalog in \texttt{FITS} (Flexible Image Transport System)
            format that contains the coordinates, magnitudes, and model parameters of 
            synthetic objects.  
            The positions of synthetic galaxies can be specified via the input catalog, 
            they can be distributed randomly over a given region, or on an 
            evenly-spaced grid (the grid option is useful in that it avoids blends 
            between Synthetic objects, e.g. see Murata\etal in prep.).

@@

        \item With the input catalog and test data information in hand, \synpipe{} 
            injects synthetic objects into single-exposure images (step
            \texttt{addFakes.py}; see the middle panel of Fig \ref{fig:flowchart}).
            \synpipe{} goes through individual CCD images that belong to 
            each \visit{} and decides which synthetic objects from the input catalog 
            need to be injected. 
            \synpipe{} uses the \textit{initial astrometric calibration} of each 
            single \visit{} to convert the input coordinates into locations on the 
            CCD image. 
            For each synthetic object, \synpipe{} uses the reconstructed PSF for each 
            exposure.  
            The photometric zero point from the single-\visit{} calibration is used
            to convert input magnitudes of synthetic objects into fluxes.

@@

%% ------------------------------------------------------------------------------------ %%            
        If necessary, an additional shear ($g_1$ and $g_2$) can be applied to the 
        model at this point.
        Then, \synpipe{} passes the reconstructed PSF image into a
        \galsim{}\texttt{InterpolatedImage} object and convolves it with the galaxy 
        model using \galsim{}\texttt{.Convolve}.
        After this, \synpipe{} uses the \galsim{}\texttt{.drawImage} method to redner 
        the image of the simulated galaxy.
        A component with a high \ser{} index ($n_{\mathrm{Ser}} > 4$) often requires 
        a large image size to cover all of its flux, so \synpipe{} allows the user 
        to truncate the model at a given radius (N$\times R_{\mathrm{e}}$).
        \synpipe{} will also give warning information when a component with $
        n_{\mathrm{Ser}} > 6$ is encountered because it takes much longer to achieve
        accurate PSF convolution in such a model; and the result is, therefore,
        not a realistic galaxy model.

        \item After \synpipe{} creates the image of the simulated galaxy, it then 
            shifts the image according to is location on the CCD on a pixel-by-pixel 
            basis; \synpipe{} also crops the images when necessary.
            When that task is complete, \synpipe{} adds appropriate noise to the 
            image based on the pixels it covers and the calibration information of 
            the detector, and creates a variance image that reflects the influence 
            of the synthetic object.
            Since each HSC CCD consists of 4 amplifiers, each with different 
            characteristics, \synpipe{} carefully creates a corresponding gain map 
            to provide accurate noise level.
            The added noise only accounts for the additional flux in the synthetic 
            object, and will thus be subdominant compared to the noise already in 
            the image for faint objects.

@@

    Using the infrastructure provided by \hscpipe{}, the user can easily perform these 
    steps in parallel, and the overall efficiency of \synpipe{} is similar to
    the real data reduction process. 
    The data and catalog volumes is also similar to the original data reduction process.

    The final step is to identify Synthetic galaxies in the output catalogs. 
    \synpipe{} can match the input catalog to the output catalog (this contains a mix 
    of real and synthetic galaxies) using a matching radius specified by the user.
    Each unmatched object from the input catalog is also passed to the results catalog
    with a unique label.
    In the case of multiple matches, the user can choose between returning only the
    closest match or returning all objects within the matching radius (the nearest one
    is labeled as such).
%% ------------------------------------------------------------------------------------ %%

@@

%% ------------------------------------------------------------------------------------ %%
\subsection{Data}
    We use \tract{}$=9699$ in the VVDS field (median FWHM$=0.449$\asec{} in $i$ band;
    in this paper referred to as \texttt{goodSeeing}) and
    \tract{}$=8764$ in the XMM-LSS field (median FWHM$=0.700$\asec{};
    \texttt{badSeeing}).
    \tract{}$=9699$ also has better seeing in both $r$ and $z$ bands than the median 
    conditions in those bands.
    The seeing conditions in $g$ and $y$ bands are very similar for both
    \tract{}$=9699$ and \tract{}$=8764$.
    For both \tracts{}, the \coadd{} image in each band includes data from
    20--40 \visits{}\footnote{The large number of \visits{} here is due to large 
    dither pattern. Please see Fig \ref{fig:flowchart}.}.  
    These two \tract{} are selected because they are not at the edge of a field, do 
    not contain any extremely bright ($i<12$ mag) saturated stars, and are 
    representative of both "good" and "bad" seeing\footnote{Here, ``bad'' seeing only 
    suggests that it is worse than the median seeing condition by HSC standard.} 
    conditions in the $i$ band.
%% ------------------------------------------------------------------------------------ %%

@@

    We use a synthetic star sample built using data from the HST/ACS catalog of 
    \citet{Leauthaud2007}.  
    We first select stars from the \citet{Leauthaud2007} catalog (this classification 
    is reliable down to $I_{\mathrm{F814W}}{\sim}25.2$ mag). 
    We then match this sample against our HSC UltraDeep--COSMOS data (which reaches 
    ${\sim}27.2$ mag in $i$ band). 
    After applying basic quality cuts (see \ref{app:qc}), this procedure provides us 
    with five-band HSC PSF photometry for 14,472 stars.
    To increase the sample size, we re-sample the five-band magnitude distributions 
    using the \texttt{astroML} (\citealt{astroml}) implementation of the 
    \texttt{extreme-deconvolution} algorithm developed by \citet{Bovy2011}
    \footnote{\texttt{https://github.com/jobovy/extreme-deconvolution/}} to
    generate 100,000 synthetic stars. 
    Fig \ref{fig:star_sample} shows the magnitude and color distributions of real 
    COSMOS stars compared to the resampled distributions in all five-bands and in 
    color-color space.

    For synthetic galaxies, we use single-\ser{} galaxy models of COSMOS galaxies with 
    $I_{F814W} \leq 25.2$ mag. 
    These are models from \citet{Lackner2012} applied to \textit{HST}/ACS images and 
    are included in \galsim{}. 
    Each model is described by a total of five parameters: magnitude, effective radius,
    \ser{} index, axis ratio, and position angle.
    Although the single-\ser{} model is not always the best choice for describing 
    galaxies, its flexibility enables it to reasonably describe the flux distribution 
    of most galaxies.
    Also, its simplicity makes it easy for us to diagnose potential problems with the 
    \cmodel{} photometry.
    We exclude a tiny fraction (${\sim}4$\%) of ill-behaved models that have a very 
    high \ser{} index ($n_{\mathrm{Ser}} > 6.0$) or a very low central surface 
    brightness ($\mu_{i} < 24.0$\sb).
    We further match this COSMOS sample with the HSC UltraDeep--COSMOS data.
    After removing objects with problematic photometry (see Appendix \ref{app:qc} for 
    details), we obtain a sample of 58,210 synthetic galaxies with realistic 
    five-band HSC \cmodel{} photometry.
    As shown in Fig \ref{fig:galaxy_sample}, the majority of these galaxies are
    faint ($i<24.0$ mag) and barely resolved ($R_{\mathrm{e}}< 1.0$\asec).
    This sample is appropriate to test \hscpipe{}'s general photometric behaviors, 
    but the sample lacks relatively bright galaxies ($i<20.5$ mag).

    We spatially distribute synthetic stars and galaxies randomly in our two selected 
    \tracts{}. 
    For stars, we use a number density of 1,000 per CCD. 
    For galaxies, we use a number density of 500 per CCD. 
    Given the magnitude distributions of these objects, these numbers are high enough 
    to ensure large sample of useful synthetic objects for the test, and will also not
    create unrealistic crowded images.

@@

    We use HSC DR1 data stored at Kavli Institute for the Physics and Mathematics of 
    the Universe (Kavli IPMU) for these tests. 
    Using 108 cores, the \texttt{addFakes.py} step takes ${\sim}1.5$ hours per 
    \tract{} in each band for stars. 
    The same process takes longer for galaxy tests (${\sim}3.0$ hours per
    \tract{} in each band) because the \galsim{} simulation is more time consuming.
    The \texttt{stack.py} step and \texttt{multiBand.py} step together take 
    ${\sim}3.5$ hours per \tract{}.

    We match the results with the input catalogs using a 2--pixel matching radius
    to generate output catalogs for \forced{} and \unforced{} photometry
    in each band. 
    For our tests, we reject synthetic objects locate within 2 pixels 
    of a real HSC object. 
    These ``ambiguously blended'' (e.g. \citealt{Dawson2016}) objects are extreme 
    blends in which multiple objects are detected as a single object, and are not 
    useful for photometric tests.

@@

    Here we discuss \hscpipe{}'s general performance, assessed mainly through \forced{} 
    PSF photometry (for stars) and (\cmodel{}) photometry (for galaxies).
    Although \hscpipe{} does provide other options (e.g., aperture and Kron photometry),
    these two are the only options that consider the effects of the PSF in different 
    bands and are consistent across all bands in terms of position and shape.

@@

    In \hscpipe{}, the PSF magnitude is derived using a matched-filter method that
    depends on the best-fit PSF model and uses third-order Lanczos interpolation to 
    shift the PSF model.
    The error of the PSF magnitude from \hscpipe{} only considers the per-pixel noise,
    and does not include centroid uncertainties.
    \hscpipe{} also estimates aperture correction for PSF photometry.
    For more details please see Bosch\etal (in prep.).

@@

    Next, we exclude stars that are misidentified by \hscpipe{} as extended
    objects using the \texttt{classification.extendedness} parameter in each band.
    The fraction of misclassification is ${\sim}10$--$20$\%, and clearly depends on
    seeing conditions.
    For the same reason, the $g$-band has the highest fraction of misclassification,
    while $i$-band is recommended for selecting point sources.
    The star--galaxy separation issue is discussed more in Section \ref{ssec:sg}.

    In the following comparisons, we show the results from the
    \texttt{goodSeeing} and \texttt{badSeeing} \tracts{} separately because it
    is important to understand how the seeing condition affects photometric accuracy 
    as well as to test whether \hscpipe{} can deliver unbiased photometry under 
    different seeing conditions.

    Besides seeing, the degree to which an object is blended with other objects 
    is another factor that influences photometric accuracy. 
    To quantify the impact of blending effects, we use the ``blendedness'' 
    parameter, $b$. 
    This parameter describes how much any given object overlaps with neighboring 
    objects (see Appendix \ref{app:defineb} and Murata\etal in prep.). 
    We divide our sample according to the blendedness parameter. 
    A value of $b=0$ corresponds to isolated objects whereas objects with $b=1$ 
    complete overlap with other objects. 
    Here we use $b>0.05$ to define highly blended stars (${\sim}4$--$9$\% are blended
    according to this criterion).
    The fraction of highly blended stars slightly increases in the bad seeing \tract{}. 
    The impact of blendedness will be discussed more in Section \ref{ssec:blendedness}.

@@

    %% C %%
    \bibitem[Calabretta \& Greisen(2002b)]{WCS2} Calabretta, M.~R., \& Greisen, E.~W.\
             2002, \aap, 395, 1077

    \bibitem[Chang et al.(2015)]{Chang2015} Chang, C., Busha, M.~T., Wechsler, R.~H.,
             et al.\ 2015, \apj, 801, 73

    %% D %%
    \bibitem[Dawson et al.(2016)]{Dawson2016} Dawson, W.~A., Schneider, M.~D., 
             Tyson, J.~A., \& Jee, M.~J.\ 2016, \apj, 816, 11

    \bibitem[Dressler et al.(2012)]{Dressler2012} Dressler, A., Spergel, D., Mountain,
             M., et al.\ 2012, arXiv:1210.7809

@@
